---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

library(kphon)
```

# kphon

Helper functions for the Kaytetye Phonological Database (kphon).

# Installation
```{r install step, eval=FALSE}
# If you don't have the devtools package, then do 'install.packages("devtools")' first
devtools::install_github("fauxneticien/kphon")
```

# Examples

Create standardised export CSV and wav files for kphon, which are of the form:

```{r Show kphon CSV, echo=FALSE}
knitr::kable(readr::read_csv("orthography,gl_annotated,transcription,annotator,kphon_export,annotation_source,xmin,xmax
atnyemilparele,south.wind-ERG,ɐcɲejmejlpɐɻələ,MP,atnyemilparele_AR_cf125a3a.wav,southwindpassage_AR_MP.TextGrid,2.1256905074478105,3.140994243207525
kape,and,kɐpə,MP,kape_AR_29cf0b23.wav,southwindpassage_AR_MP.TextGrid,3.2237406016823074,3.606705482796561
aherrkele,sun-ERG,ɐːɾkələ,MP,aherrkele_AR_4bcff7be.wav,southwindpassage_AR_MP.TextGrid,3.606705482796561,4.07538582324225
tyampe,too,cɐmpə,MP,tyampe_AR_95ac9af5.wav,southwindpassage_AR_MP.TextGrid,4.07538582324225,4.511354649921463
ahenge,fight,ɐːŋə,MP,ahenge_AR_e67432f2.wav,southwindpassage_AR_MP.TextGrid,4.660752408376567,4.950564710693483"))
```


**Step 0:** Load kphon library and set working directory

```{r demo step, eval= FALSE}
library(kphon)

setwd("~/sandboxes/kphon-v2/annotations/southwind/")
```

**Step 1:** Provide TextGrid file(s) to process, with an Input/Output mapping

```{r process raw textgrids, eval=FALSE}
# Step 1: Provide input TextGrid file(s)
#         and an input-output 'io' mapping for TextGrid tiers to columns
step1 <-
    textgrid_to_kphon(
        input_files = "southwindpassage_AR_MP.TextGrid",
        io_mapping = list(
            orthography   = "word",
            gl_annotated  = "gloss",
            transcription = "phon"
        )
    )
```

**Step 2:** Derive (however appropriate) speaker and annotator columns, then feed into hash function to derive file name for export

```{r make name cols, eval=FALSE}
step2 <-
    step1 %>%
    mutate(
        speaker      = str_extract(string = annotation_source, pattern = "_[A-Z]+_") %>% str_remove_all("_"),
        annotator    = str_extract(string = annotation_source, pattern = "[A-Z]+\\.TextGrid") %>% str_remove("\\.TextGrid"),
        input_wav    = str_replace(string = annotation_source, pattern = "_[A-Z]+.\\TextGrid", ".wav"),
        output_wav   = hash_kphon_wav(orthography, speaker, annotation_source, xmin)
    )

```

**Step 3:** After you've verified data extracted correctly in Step 1 and Step 2, write .wav and .csv files

```{r write_kphon, eval=FALSE}
write_kphon_exports(kphon_df = step2, path = ".", overwrite = TRUE)
```

